---
title: "Mapping Quandl Data"
output: html_notebook
---
In previous posts, we have built a [map to access global ETFs](https://rviews.rstudio.com/2016/12/16/reproducible-finance-with-r-a-shiny-etf-map/) and a simple [Shiny app to import and forecast commodities data from Quandl](https://rviews.rstudio.com/2017/04/21/a-shiny-app-for-importing-and-forecasting-commodities-prices-from-quandl/). Today, we will begin a project that combines those previous apps. Our end goal is to build an interactive map to access macroeconomic data via Quandl, allowing the user to choose an economic indicator and click on a country to access that indicator's time series. For example, if a user wants to import and graph GDP-per-capita data from China, that user will be able to select `GDP per Capita` from a drop down and then click on China to display the time series.  As usual, we'll start with a Notebook for map building, data import and test visualizations.  

Careful readers will note that we are not simply reusing code from those previous projects and that is because we are going to introduce some new packages to our toolkit.  Specifically: 

- Instead of downloading a spatial dataframe from [Natuarl Earth](http://www.naturalearthdata.com/), we will use the new r package [rnaturalearth](https://cran.r-project.org/web/packages/rnaturalearth/rnaturalearth.pdf) and it's function `ne_countries()` to import our geographic data. 

- Relatedly, we are not going work with a "spatial" dataframe but with a simple features dataframe of class `sf`. Learn more [here](https://cran.r-project.org/web/packages/sf/vignettes/sf1.html).

- Finally, when we test our data import of several macroeconomic data sets, we will use the `map()` function from the [purrr](https://cran.r-project.org/web/packages/purrr/purrr.pdf) instead of relying on `lapply()`, as we did previously when passing ticker symbols to `getSymbols()`. If you're like me, you won't be sad to be done with `lapply()`.

Let's get to it!

We are going to be working with a macroeconomic data source from the World Bank called World Development Indicators [WDI](http://data.worldbank.org/products/wdi).  The Quandl code for WDI is `WWDI` and thus we'll append `WWDI/` to each data set call.

One difference between our previous work with Quandl and today, is that previously we imported and visualized global time series, such as the price of oil, gold or copper. By global, I mean these time series were not associated with a particular country. That simplified our job a bit because we did not need to supply a country code to Quandl. Today, we do need to include a country code.

Let's start with a simple example and import GDP-per-capita data for China, whose country code is `CHN`.  We will pass `WWDI/CHN_NY_GDP_PCAP_KN` to Quandl, which is the WDI code `WWDI/` appended to the China country code `CHN` appended to the GDP-per-capita code `_NY_GDP_PCAP_KN`. 

```{r China GDPPC, message = FALSE}
library(Quandl)
library(tidyverse)
China_GDPPC <- Quandl("WWDI/CHN_NY_GDP_PCAP_KN", type = 'xts') %>% 
  # Add a nice column name
  `colnames<-`("GDP Per Capita")

tail(China_GDPPC, n = 6)
```
That looks pretty good, but note that the data goes only to December of 2017 - not ideal but fine for our illustrative purposes. 

Our eventual Shiny app, though, will give the user a choice among many economic time series and we want to test the import and visualiation of all those choices. In addition to GDP-per-capita, let's test GDP-per-capita growth, real interest rate, exchange rate, CPI and labor force participation rate. Those offer a nice snapshot of a country's economy.

To run our test in the code chunk below, we will first build a vector of dataset codes. Note that each of them starts with `WWDI/`, then `CHN`, and then the WDI code. Next we will pipe that vector to Quandl using the `map()` function. `map()` takes a function a function, in this case the `Quandl()` function and applies it to the elements in a vector, in this case our vector of dataset codes. We still get to specify `type = "xts"`. 

If stopped our pipe there, the output would be a list of 6 time series and that would suit our testing purposes but I don't love having to look at a list of 6. Thus, we'll keep on piping and use the `reduce()` function from the purrr package to `merge()` the 6 lists into one xts object. That is much easier to deal with and easier to pass to dygraphs for our test visualization. The last piped function will invoke `colnames<-` to clean up the column names.  

```{r Many Econ Indicators}

# Create a vector of economic indicators that can be passed to Quandl via map().
# Include names and values for easy naming of xts columns.
econIndicators <- c("GDP Per Capita" = "WWDI/CHN_NY_GDP_PCAP_KN",
                  "GDP Per Capita Growth" = "WWDI/CHN_NY_GDP_PCAP_KD_ZG",
                  "Real Interest Rate" = "WWDI/CHN_FR_INR_RINR",
                  "Exchange Rate" = "WWDI/CHN_PX_REX_REER",
                  "CPI" = "WWDI/CHN_FP_CPI_TOTL_ZG",
                  "Labor Force Part. Rate" = "WWDI/CHN_SL_TLF_ACTI_ZS")
# You might want to supply an api key
Quandl.api_key("d9EidiiDWoFESfdk5nPy")

China_all_indicators <- 
  # Start with the vector of Quandl codes
  econIndicators %>% 
  # Pass them to Quandl via map(). 
  map(Quandl, type = "xts") %>% 
  # Use the reduce() function to combine them into one xts objects.
  reduce(merge) %>% 
  # Use the names from the original vector to set nicer column names.
  `colnames<-`(names(econIndicators))

tail(China_all_indicators, n = 6)
```

```{r}
# Include names and values for easy naming of xts columns.
econIndicators2 <- c("GDP Per Capita" = "_NY_GDP_PCAP_KN",
                  "GDP Per Capita Growth" = "_NY_GDP_PCAP_KD_ZG",
                  "Real Interest Rate" = "_FR_INR_RINR",
                  "Exchange Rate" = "_PX_REX_REER",
                  "CPI" = "_FP_CPI_TOTL_ZG",
                  "Labor Force Part. Rate" = "_SL_TLF_ACTI_ZS")

China_all_indicators2 <- 
  # Start with the vector of Quandl codes
  econIndicators2 %>%
  map_chr(~paste("WWDI/", "CHN", ., sep="")) %>% 
  # Pass them to Quandl via map().
  map(Quandl, type = "xts") %>% 
  
  # Use the reduce() function to combine them into one xts objects.
  reduce(merge) %>% 
  # Use the names from the original vector to set nicer column names.
  `colnames<-`(names(econIndicators2))
tail(China_all_indicators2, n=6)
econIndicators
```


That seems to have done the job, and note that both CPI and Exchange have data up to December of 2016. Much better and a sign that there will be date variation amongst countries and indicators. That won't be crucial today but it could be if we needed consistent dates for all of our datasets. 

As a final step, let's use `dygraph()` to visualize these time series to make sure that will function when we move to Shiny.

```{r Dygraph}
library(dygraphs)
dygraph(China_all_indicators$`GDP Per Capita`, main = "GDP Per Capita")
dygraph(China_all_indicators$`GDP Per Capita Growth`, main = "GDP Per Capita Growth")
dygraph(China_all_indicators$`Real Interest Rate`, main = "Real Interest Rate")
dygraph(China_all_indicators$`Exchange Rate`, main = "Exchange Rate")
dygraph(China_all_indicators$`CPI`, main = "CPI")
dygraph(China_all_indicators$`Labor Force Part. Rate`, main = "Labor Force Part. Rate")
```

```{r}
library(highcharter)
test_hc <- China_all_indicators2[,1]
hc <- highchart(type = "stock") %>% 
  hc_title(text = "Charting Econ Data From Quandl") %>% 
  hc_subtitle(text = "Data extracted using Quandl package") %>% 
  hc_add_series(China_all_indicators2[,2], name = names(China_all_indicators2[,2]), id = "test") %>% 
  hc_add_series(China_all_indicators2[,3], name = names(China_all_indicators2[,3]), id = "test1") 
hc
```


GDP-per-capita has been on a steady march higher, labor force participation plummeted in the 90's and 2000's, probably as the labor market became more market-oriented and real interest rates look quite choppy. Plenty to explore and it signals that our Shiny app might be a nice exploratory tool. 

For now, it's on to map-building! 

In previous posts, we have started map-building by grabbing a spatial object from the internet. Today, we will use the new rnaturalearth package and it's fantastic function `ne_countries()`. We will specify `type = "countries"` and `returnclass = 'sf'`.  We could have set `returnclass = 'sp'` if we wanted to work with a spatial dataframe but I have been migrating over to sf package and simple features data frames for mapping projects. Lots more to learn here but in general it seems to use more tidy concepts than the sp package.  

```{r Import SF Object, message = FALSE, warning = FALSE}
library(rnaturalearth)
library(rgeos)
library(sf)

world <- ne_countries(type = "countries",  returnclass='sf')

# Take a peek at the name, gdp_md_est column and economy columns. 

world %>% 
  select(name, gdp_md_est, economy, geometry) %>% 
  head()
```

This output looks similar to an `sp` object, but it's not identical. In particalur, note the `geometry` column which contains the multipolygon latitude and longitude coordinates. It's a matter of personal preference, but I find this `sf` object more intuitive in terms of how the geometry corresponds with the rest of the data. 

Now we'll create our shading and popups. Note that this code is identical to when we were working with spatial dataframes.

```{r Build Map, message = FALSE, warning = FALSE}
library(leaflet)
##create shading by GDP
gdpPal <- colorQuantile("Blues", world$gdp_md_est, n = 20)

##create popup country name and economic stage
popup <- paste0("<strong>Country: </strong>", 
                world$name, 
                "<br><strong>Market Stage: </strong>", 
                world$income_grp)
```

Now, it's time to build our leaflet map. This code is nigh identical to our code from previous projects, but with one major change.  We need to pass in a country code to Quandl, not a country name. Thus, we don't want to set `layerId = name` but we also don't need to add anything special to our `world` object. It already contains a column called `iso_a3` and that column contains the country codes used by Quandl. Fantastic! We just need to set `layerId = ~iso_a3` and we will be able to access country codes in our Shiny app. This is fortunate but it also highlihgts a good reason to spend some time studying the `world` object.  It might contain other data that will be useful in future projects. 

```{r Leaflet Map Build, warning = FALSE, message = FALSE}

leaf_world <- leaflet(world) %>%
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng =  20, lat =  15, zoom = 2) %>%
      addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = .7, 
      # Note the layer ID. Not a country name! It's a country code! 
      color = ~gdpPal(gdp_md_est), layerId = ~iso_a3, popup = popup)

leaf_world
```

Because Quandl uses the [ISO 3-letter code](https://blog.quandl.com/api-for-economic-data) to identify a country and because the rnaturalearth `sf` object already contained a column with iso_a3 country codes, building our map wasn't very hard.  It's worth pointing out that in various projects, we have now built maps with 3 different layerIDs. In this project, we built one map with `layerID = name`, using the built-in `name` column in the rnaturalearth object, and we built one map with `layerID = tickers`, using a column of tickers that we added ourselves with the `merge()` function. Now, today, we have a map with `layerID = iso_a3`, using the built-in column of country codes. We are building up a nice portfolio of maps with different layerIDs should we ever want or need them in the future. 

That's all for today. We'll save that `leaflet_world` object for use in our Shiny app! See you next time. 


```{r}
library(here)
# Save
save(leaf_world,  file = here("map-data", 'gdp-colored-map.RData'))
```




